\thetitle

### 2020-11-08 sun_phases/spin_qudit_tomo.tex
tighten upper bound on reconstruction error

Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the transition operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
We can expand the transition operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
\begin{align}
  T_{i\ell} = \sum_\mu t_{\ell\mu} \Pi_{i\mu},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu},
\end{align}
and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
The covariances between statistical errors in the transition operator basis are then
\begin{align}
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell'}}
  = \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \sp{\delta_{\mu\mu'} p_\mu^{(i)} - p_\mu^{(i)} p_{\mu'}^{(i)}}.
\end{align}
In order to establish an upper bound on reconstruction error for an unknown quantum state, we need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
To this end, for any multinomial probability distribution $p\equiv\p{p_1,p_2,\cdots,p_d}$ we define the function
\begin{align}
  f_\ell\p{p} \equiv \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell\mu'}
  \p{\delta_{\mu\mu'} p_\mu - p_\mu p_{\mu'}}
  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2
  \le \sum_\mu t_{\ell\mu}^2 p_\mu
  \le \sum_\mu t_{\ell\mu}^2 = 1,
\end{align}
which implies that
\begin{align}
  \bbk{\epsilon_{i\ell}^2} = \f1n f_\ell\p{p^{(i)}} \le \f1n.
  \label{eq:variance_bound}
\end{align}

@@

Equipped with the bound on variances in Eq.~\eqref{eq:variance_bound}, we again consider the block-diagonalized form of the measurement matrix derived in Section \ref{sec:bound}:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0} \op{m}{\v v}.
\end{align}
The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound} of Section \ref{sec:bound} without the assumption in Eq.~\eqref{eq:error_assumption}, finding that
\begin{align}
  \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  = \sum_{i,i'} q_{\ell m i}^V \p{q_{\ell m i'}^V}^*
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_i \abs{q_{\ell m i}^V}^2
  \bbk{\epsilon_{i\ell}^2}
  \le \f1n \sum_i \abs{q_{\ell m i}^V}^2
  = \f1n \times \f1{\p{M_{\ell m}^V}^2}.
\end{align}
Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of the unknown state $\rho$ that is accurate to within a mean squared distance
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} = \sum_{\ell,k}
  \Bbk{\abs{\tilde\rho_{\ell k}^V-\rho_{\ell k}^V}^2}
  \le \f{\S_V^2}{n},
\end{align}
where $\S_V$ is the same error scale as in Section \ref{sec:bound} of the main text.

>>>>

Rather than the statistical errors $\epsilon_{i\mu} \equiv \tilde\Pi_{i\mu} - \bk{\Pi_{i\mu}}_\rho$ in the estimates $\tilde\Pi_{i\mu}$ of the projectors $\Pi_{i\mu}$, we now consider the statistical errors $\epsilon_{i\ell} \equiv \tilde T_{i\ell} - \bk{T_{i\ell}}_\rho$ in the estimates $\tilde T_{i\ell}$ of the transition operators $T_{i\ell} \equiv T_{\v v_i\ell,0}$.
We can expand the transition operators $T_{i\ell}$ as a sum over projectors $\Pi_{i\mu}$ as
\begin{align}
  T_{i\ell} = \sum_\mu t_{\ell\mu} \Pi_{i\mu},
  &&
  t_{\ell\mu} \equiv \bk{\mu|T_{\ell,0}|\mu},
\end{align}
and likewise $\tilde T_{i\ell} = \sum_\mu t_{\ell\mu} \tilde\Pi_{i\mu}$.
The covariances between statistical errors in the transition operator basis are then
\begin{align}
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell'}}
  = \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \bbk{\epsilon_{i\mu} \epsilon_{i'\mu'}}
  = \delta_{ii'} \times \f1n \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell'\mu'}
  \sp{\delta_{\mu\mu'} p_\mu^{(i)} - p_\mu^{(i)} p_{\mu'}^{(i)}}.
\end{align}
In order to establish an upper bound on reconstruction error for an unknown quantum state, we need an upper bound on the variance $\bbk{\epsilon_{i\ell}^2}$ in particular.
To this end, for any multinomial probability distribution $p\equiv\p{p_1,p_2,\cdots,p_d}$ we define the function
\begin{align}
  g_\ell\p{p} \equiv \sum_{\mu,\mu'} t_{\ell\mu} t_{\ell\mu'}
  \p{\delta_{\mu\mu'} p_\mu - p_\mu p_{\mu'}}
  = \sum_\mu t_{\ell\mu}^2 p_\mu - \p{\sum_\mu t_{\ell\mu} p_\mu}^2.
\end{align}
Ideally, we would maximize this function over all valid probability distributions, but doing so is rather difficult.
Instead, we can find a loose upper bound by ignoring the negative terms:
\begin{align}
  g_\ell\p{p} \le \sum_\mu t_{\ell\mu}^2 p_\mu
  \le \max_\mu t_{\ell\mu}^2
  = \f{2\ell+1}{d} \max_\mu \bk{s\mu;\ell,0|s\mu}^2
  \le \f{2\ell+1}{d},
\end{align}
which implies that
\begin{align}
  \bbk{\epsilon_{i\ell}^2} = \f1n g_\ell\p{p^{(i)}} \le \f{2\ell+1}{nd}.
  \label{eq:variance_bound}
\end{align}
Note that $t_{0,\mu}=1/\sqrt{d}$, so $f_0\p{p}=0$ and $\bbk{\epsilon_{i,0}^2}=0$.

@@

Equipped with the bound on variances in Eq.~\eqref{eq:variance_bound}, we again consider the block-diagonalized form of the measurement matrix derived in Section \ref{sec:bound}:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
  &&
  M_{V\ell} = \sum_{m,\v v} D^\ell_{m,0} \op{m}{\v v}.
\end{align}
The block diagonal structure of this matrix implies that, rather than an integer index $k\in\set{1,2,\cdots,d^2}$, we can identify the singular values and right singular vectors of $UM_V$ (and equivalently $M_V$) by a block index $\ell\in\set{0,1,\cdots,d-1}$ and a sub-index $m=\set{-\ell,-\ell+1,\cdots,\ell}$.
Making the replacement $k\to\p{\ell,m}$, we can thus repeat the derivations through Eq.~\eqref{eq:bound} of Section \ref{sec:bound} without the assumption of Eq.~\eqref{eq:error_assumption}, finding that
\begin{align}
  \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  = \sum_{i,i'} q_{\ell m i}^V \p{q_{\ell m i'}^V}^*
  \bbk{\epsilon_{i\ell} \epsilon_{i'\ell}}
  = \sum_i \abs{q_{\ell m i}^V}^2
  \bbk{\epsilon_{i\ell}^2}
  \le \f{2\ell+1}{nd} \p{1-\delta_{\ell,0}}.
\end{align}
Altogether, we thus find that $n$ measurements of spin projection along every axis $\v v\in V$ allows us to construct an empirical estimate $\tilde\rho_V$ of an unknown state $\rho$ that is accurate to within a mean squared distance
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \sum_{\ell,m} \Bbk{\abs{\tilde\rho_{\ell m}^V-\rho_{\ell m}^V}^2}
  \le \f1{nd} \sum_{\ell>0} \p{2\ell+1} \S_{V\ell}^2,
\end{align}
where
\begin{align}
  \S_{V\ell} \equiv \norm{M_{V\ell}^{-1}}
  = \sqrt{\sum_m \p{M^V_{\ell m}}^{-2}}
\end{align}
is the (Hilbert-Schmidt, or Frobenius) norm of the right inverse of $M_{V\ell}$.
